---
title: "Assignment 2"
output: pdf_document
date: "2022-11-03"
author: B. Nunez
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(e1071)
```

## R Markdown

NOTE that tidyverse and e1071 packages must be installed.

Collecting the data:

```{r data, echo=FALSE}
df <- read_csv("NYPD_Shooting_Incident_Data__Historic_.csv")
```

Cleaning up and formatting the data: remove rows where the age and race values are empty, then summarize the data.

```{r cars, echo=FALSE}
summary(df)
data_by_age <- df$PERP_AGE_GROUP
data_by_race <- df$PERP_RACE
data_by_age <- data_by_age[!is.na(data_by_age)]
age_table <- table(data_by_age)
age_table <- age_table[!(names(age_table) %in% c("940","UNKNOWN"))]
data_by_race <- data_by_race[!is.na(data_by_race)]
race_table <- table(data_by_race)
dfx <- df[c("PERP_RACE", "VIC_RACE")]
dfx$PERP_RACE <- as.factor(dfx$PERP_RACE)
dfx$VIC_RACE <- as.factor(dfx$VIC_RACE)
df <- na.omit(df)
```

## Including Plots

Visualizations:

```{r barplot, echo=FALSE}
barplot(age_table, main="Perpetrator Age Ranges", xlab="Age Range", ylab="Count")
barplot(race_table, main="Perpetrator Race", xlab="Race", ylab="Count", cex.names=0.5)
```

Note that erroneous or unknown value data points were removed in the visualizations. We can see that the perpetrator in the crime is disproportionately inclined to one race and age.

We create a model to predict victim race based on perpetrator race:

```{r model, echo=FALSE}
set.seed(100)
dfx <- dfx[sample(1:nrow(dfx)), ]
split <- round(nrow(dfx) * 0.80)
train <- dfx[1:split, ]
test <- dfx[(split + 1):nrow(dfx), ]
nb <- naiveBayes(PERP_RACE ~ VIC_RACE, data = train)
pred <- predict(nb, test)
tab1 <- table(pred, test$VIC_RACE)
print(tab1)
sum(diag(tab1)) / sum(tab1)
```

This shows that the model predicts the victim race based on the perpetrator race with about 74% accuracy.

Sources of bias: Possibly the significant omission of unknown and erroneous data points that may have changed the model in some way, were their true values to be known. 
Also, there too many black perpetrator / victim data points, so that the model is biased in such a way that it is difficult to tell if there is a statistically significant relationship between races in general, versus only black perpetrator / victim data points.

